An Improved Metric for the Analysis of Swarms

### 2018-04-18 paper.tex
Minor abstract change and new flood graphics

\begin{abstract}
An important criterion in the assessment of swarm coordination algorithms is
the extent to which they cause unnecessary movement, i.e. movement which
occurs as a side affect of the algorithm rather than the the intended
function. This paper examines the widely used distance metric as a mechanism
to measure the internal movement of agents and introduces a new
\emph{cohesion/repulsion metric}.  Both metrics are based upon the resultant
internal movement of a swarm identified by analysing the changes in the
inter-agent interactions. The two metrics differ in their approach to
identifying the changes. The distance metric uses variations in the
inter-agent distances and the new metric uses the resultant magnitudes from
the agents' \emph{inter-agent vector magnitudes}  which we claim offers a
number of advantages over the distance metric.  Both metrics allow a
comparison of the effects of different swarming algorithms on a swarm's
structure to identify the `level' of change in a swarm's internal structure
which is an indication of the amount of unnecessary movement an algorithm
creates.  In addition the \emph{cohesion/repulsion} metric also identifies a
swarm as expanding or cohesive.
\end{abstract}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability1}
{Internal movement cohesion (no repulsion).\label{methods:Stability1}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability2}
{Internal movement repulsion.\label{methods:Stability2}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability3}
{Internal movement cohesion and repulsion.\label{methods:Stability3}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability4}
{Internal movement equilibrium.\label{methods:Stability4}}

@@

The field effects are incremented in turn, cohesion range followed by repulsion range. After each repulsion field change the swarm is allowed to redistribute itself and stabilise. Table~\ref{tab:FillSequence} shows the field effect settings that are used for the simulation. The field effects are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm. Figure \ref{emerge:FLOOD1} shows the initial swarm deployment and Figure \ref{emerge:FLOOD6} shows the final swarm distribution for the simulation.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD1}
{Flood fill Start.\label{emerge:FLOOD1}}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FLOOD6}
{Flood fill End.\label{emerge:FLOOD6}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=16cm]{figures/FIELDFILL-MAG}
{Cohesion/Repulsion metric 0-120 seconds.\label{emerge:FIELDFILL-MAG}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=16cm]{figures/FIELDFILL-DIST}
{Distance metric 0-120 seconds.\label{emerge:FIELDFILL-DIST}}

>>>>

\begin{abstract}
An important criterion in the assessment of swarm coordination algorithms is
the extent to which they cause unnecessary movement, i.e. movement which
occurs as a side affect of the algorithm rather than the the intended
function. This paper examines the widely used distance metric as a mechanism
to measure the internal movement of agents and introduces a new
\emph{cohesion/repulsion metric}.  Both metrics are based upon the resultant
internal movement of a swarm identified by analysing the changes in the
inter-agent interactions. The two metrics differ in their approach to
identifying the changes. The distance metric uses variations in the
inter-agent distances and the new metric uses the resultant magnitudes from
the agents' \emph{inter-agent vector magnitudes}  which we claim offers a
number of advantages over the distance metric.  Both metrics allow a
comparison of the effects of different swarming algorithms on a swarm's
structure to identify the `level' of change in a swarm's internal structure
which is an indication of the amount of unnecessary movement an algorithm
creates.  In addition the \emph{cohesion/repulsion} metric also identifies a
swarm as expanding or cohesive which can be used in the analysis of a swarm 
filling a space to identify an exit condition.
\end{abstract}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability1}
{Internal movement cohesion (no repulsion).\label{methods:Stability1}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability2}
{Internal movement repulsion.\label{methods:Stability2}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability3}
{Internal movement cohesion and repulsion.\label{methods:Stability3}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability4}
{Internal movement equilibrium.\label{methods:Stability4}}

@@

The field effects are incremented in turn, cohesion range followed by repulsion range. After each repulsion field change the swarm is allowed to redistribute itself and stabilise. Table~\ref{tab:FillSequence} shows the field effect settings that are used for the simulation. The field effects are selected so as to ensure the swarm parameters have the potential to create a hexagonal swarm. Figure \ref{emerge:FLOOD0} shows the initial swarm deployment.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD0}
{Flood fill Start.\label{emerge:FLOOD0}}

Figure ~\ref{emerge:FLOOD1} shows the point a which data logging starts. The swarm at this stage has expanded into the space but is not filling it. This presents itself as a positive cohesion/repulsion metric value which indicates the swarm is cohesive and the agents are still ``pulling" together. Figure \ref{emerge:FLOOD6} shows the final swarm distribution for the simulation, at this point the cohesion/repulsion metric has fell below zero indicating the swarm is trying to expand.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD1}
{Data Capture Start.\label{emerge:FLOOD1}}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD6}
{Flood fill End.\label{emerge:FLOOD6}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=13cm]{figures/FIELDFILL-MAG}
{Cohesion/Repulsion metric 0-120 seconds.\label{emerge:FIELDFILL-MAG}}

@@

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=13cm]{figures/FIELDFILL-DIST}
{Distance metric 0-120 seconds.\label{emerge:FIELDFILL-DIST}}
